---
title: "Data 607 Assignment 2"
author: "Joseph Simone"
date: "9/10/2019"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Assingment 2 - SQL and R 

Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates,
imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the finalults
(observations) and store them in a SQL database. Load the information into an R dataframe.


## The Set Up in SQL

I aquired a user ratings of movies dataset from, https://www.kaggle.com/shubhammehta21/movie-lens-small-latest-dataset#ratings.csv, and import a small amount of the data set into MySql server.

I used the below code to create the database:

*/
CREATE DATABASE movies;
/*


*/
USE movies;

CREATE TABLE ratings
( 
userId int(250) NULL , 
movieId int(250) NULL, 
rating int(10) NULL
) ENGINE=InnoDB;


Then I imported the database using the import wizard tool,into the table 'ratings' with the database 'movies' I created




## Connection to MySQL in R

```{r}
library(RMySQL)
```


```{r}
db <- dbDriver("MySQL")
```



```{r}
mydb <- dbConnect(db, username="root", password="Yogibear1", dbname ="movies", host="localhost")
```



```{r}
sql <- "SELECT * from ratings"
```

```{r}
final <- dbGetQuery(mydb, sql)
final
```


## Beginning of Data Transformation 
I started giving userId variable's, that were repfinalenting with number, names. I started doing the same for the movieId variable.
```{r}
final$userId[final$userId == "1"] <- "Kristyn"
final$userId[final$userId == "2"] <- "Keira"
final$userId[final$userId == "3"] <- "Dom"
final$userId[final$userId == "4"] <- "Sharon"
final$userId[final$userId == "5"] <- "Vito"
final$userId[final$userId == "6"] <- "Justin"
 
 
final$movieId[final$movieId == "1"] <- "Once Upon a Time in Hollywood"
final$movieId[final$movieId == "3"] <- "The Lion King"
final$movieId[final$movieId == "6"] <- "Fast & Furious: Hobbs & Shaw"
final$movieId[final$movieId == "7"] <- "Dora the Explorer"
final$movieId[final$movieId == "50"] <- "It: Chapter 2"
final$movieId[final$movieId == "70"] <- "Angel has"
```