---
title: "Data 607 Assignment 5"
author: "Joseph Simone"
date: "9/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyr)
library(dplyr)
library(png)
library(ggplot2)
```

Libaries used:
library(knitr)
library(png)
library(tidyr)
library(dplyr)

## Tidying and transforming Data

The chart below describes arrival delays for two airlines across five destinations. 

Your task is to:

(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above.

You’re encouraged to use a “wide” structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy
and transform your data.

(3) Perform analysis to compare the arrival delays for the two airlines.

(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative
descriptions of your data cleanup work, analysis, and conclusions.


```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/DATA Acquisition and Management/4.png"
include_graphics(imgage)
```

## DataFrame Creation

```{r}
flight_times <- rbind(c(NA, NA, "Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle"),
             c("ALASKA", "On Time", 497, 221, 212, 503, 1841),
             c(NA, "Delayed", 62, 12, 20, 102, 305),
             c("AM WEST", "On Time", 694, 4840, 383, 320, 201),
             c(NA, "Delayed", 117, 415, 65, 129, 61))
```


## Local .cvs Export

```{r}
write.csv(flight_times,'C:/Users/jpsim/Documents/DATA Acquisition and Management/flight_times.csv', row.names = FALSE)
```

## GitHub Data Load
```{r}
raw_data <- read.csv(paste0("https://raw.githubusercontent.com/josephsimone/DATA607/master/flight_times.csv"), 
                             stringsAsFactors = F); raw_data
```

## Tidying the dataset with the use of tidyr

```{r}
 clean_data <- raw_data %>%
  mutate(Airlines1 = NA., Airlines2=lag(NA.)) %>% 
  mutate(Airline = coalesce(Airlines1,Airlines2), Status = NA..1)  %>%
  gather("Destination", "Flights", 3:7) %>% 
  select(Airline:Flights) %>% 
  arrange(Airline, desc(Status), Destination); clean_data
```

## Use the dplyr package to transform the data for analysis

```{r}
 flights_aggr_df <- clean_data %>% select(Airline, Status, Flights) %>% 
      group_by(Airline, Status) %>% 
      summarise(total = sum(Flights)) %>% 
      mutate(percent           = round(total/sum(total) * 100, 2),
             percent_formatted = paste0(round(total/sum(total) * 100, 1), "%"),
             total_flights     = sum(total)) 
```


## Creation of new DataFram
```{r}
tidy_df <- as.data.frame(flights_aggr_df)
tidy_df
```



## Vizualization of Data using GGPlot2

```{r}
tidy_df %>% filter(Status == "Delayed") %>%
    ggplot( aes(x=Airline, y=percent, fill=Airline)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black", width = 0.5) +
    coord_flip() + 
    ggtitle("Overall Percentage of Flights Delayed by Airline") +
    xlab("Airline") + ylab("Percentage of Flights Delayed") +
    geom_text(aes(label=paste(percent_formatted, "\n n =", total_flights)), vjust=0.5, hjust=1.1,color="black")
```

```{r}
flights_aggr_df <- clean_data %>% select(Airline, Status, Flights, Destination) %>% 
      group_by(Airline, Status, Destination) %>% 
      summarise(total = sum(Flights)) %>% 
      mutate(percent           = round(total/sum(total) * 100, 2),
            percent_formatted = paste0(round(total/sum(total) * 100, 1), "%"),
            total_flights     = sum(total)) 
```



```{r}
flights_by_city_df <- flights_aggr_df %>%
group_by(Airline, Destination) %>% 
    mutate(percent = round(total/sum(total)*100, 2),
           percent_formatted=paste0(round(total/sum(total)*100, 1), "%")) 
```


```{r}
flights_by_city_df <- as.data.frame(flights_by_city_df)
flights_by_city_df
```


```{r}
flights_by_city_df %>% filter(Status == "Delayed") %>%
ggplot( aes(x=Destination, y=percent, fill=Airline)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black") +
    geom_text(aes(label=percent_formatted), vjust=.5, hjust=1,position= position_dodge(width=0.9),  color="black") +
        ggtitle("Percentage of Flights Delayed by Airline by City") +
    xlab("Destination") + ylab("Percentage of Flights Delayed") +
        coord_flip() 
```

