---
title: "Data 607 Project 1"
author: "Joseph Simone"
date: "9/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(reshape2)
library(downloader)
library(htmlTable)
library(dplyr)
library(stringr)
library(ggplot2)
library(DT)
library(png)
library(knitr)

```

## Chess Data File Project


In this project, you’re given a text file with chess tournament results where the information has some structure. 

Yourjob is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)

with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and 

Average Pre Chess Rating of Opponents

For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played


```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/DATA Acquisition and Management/Project1.png"
include_graphics(imgage)
```



## Data Import and Set-up

```{r}
chessdata <-'https://raw.githubusercontent.com/josephsimone/DATA607/master/tournamentinfo.txt'
file <- "tournamentinfo.txt"

downloader::download(chessdata, file)

c<- file(file, open="r")
tournament.raw <- readLines(c, warn = FALSE)
close(c)
```

```{r}
#tournament.raw
```


## Data Cleaning and Preprocessing 

```{r}
 tournament.data. <- str_replace_all(str_replace_all(tournament.raw, "->", ">>"), "-{3,}", "")

tournament.data. <- tournament.data.[tournament.data.  != ""] 

tournament.data. <- tournament.data.[-(1:2)]

tournament1 <- tournament.data.[str_detect(str_sub(tournament.data., 1, 6), "[0-9]")]

 tournament2 <- tournament.data.[str_detect(str_sub(tournament.data., 1, 6), "[A-Z]{2,2}")]
```


```{r}
head(tournament1, 5)
```


```{r}
head(tournament2, 5)
```

## Parsing into two Data-Frames
```{r}
tourny1df <- data.frame(player_num   = as.numeric(substr(tournament1, 1,  6)),
                         player_name  = str_trim(substr(tournament1, 8,  40), side="both"),
                         total_pts    = as.numeric(substr(tournament1, 42, 46)),
                         round1       = substr(tournament1, 48, 52),
                         round2       = substr(tournament1, 54, 58),
                         round3       = substr(tournament1, 60, 64),
                         round4       = substr(tournament1, 66, 70),
                         round5       = substr(tournament1, 72, 76),
                         round6       = substr(tournament1, 78, 82),
                         round7       = substr(tournament1, 84, 88), stringsAsFactors=FALSE)
```


```{r}
tourny2df <- data.frame(player_state  = str_trim(substr(tournament2, 1,  6), side="both"),
                           uscf_id       = str_extract(substr(tournament2, 8,  40), "\\d+"),
                           pre_rating    = as.numeric(str_extract(substr(tournament2, 8,  40), "(?<=R: ).\\d+(?=)")),
                           post_rating   = as.numeric(str_extract(substr(tournament2, 8,  40), "(?<=>>).\\d+(?=)")),
                           stringsAsFactors=FALSE)
```

## Joining of two newly created Data-Frames

```{r}
finaldf <- cbind(tourny1df, tourny2df)
players <- select(finaldf, player_num:total_pts, player_state:post_rating)
```


## Creation of Final Output
```{r}
matches_played <-  
      finaldf %>% select(player_num, round1:round7) %>%
      melt(id.var=c("player_num"), value.name="outcome_opp")  %>% 
     mutate(round = as.numeric(str_replace(variable, "round", "")),
              outcome = str_extract(outcome_opp, "^\\w+"),
              opponent_num = as.numeric(str_extract(outcome_opp, "\\d+$"))) %>%
      select(-c(variable, outcome_opp)) %>%
      inner_join(select(players, player_num, pre_rating), c("opponent_num" = "player_num")) %>%
      select(player_num, round, outcome, opponent_num, pre_rating) %>%
      arrange(player_num, round)
```

## Variable change
```{r}
names(matches_played)[names(matches_played) == "pre_rating"] <- "opponent_pre_rating"
```

## Calculation of Opponent's Average

```{r}
playersdf <-  
      matches_played %>%
      group_by(player_num) %>% summarise(opponents_avg_pre_rating = round(mean(opponent_pre_rating))) %>%
      inner_join(players, by="player_num") %>%
     select(player_num, player_name, player_state, total_pts,  uscf_id, pre_rating, post_rating, opponents_avg_pre_rating)


```

## Searchable Data-Frame

```{r}

#datatable(playersdf, extensions = 'Scroller', options = list(
#    deferRender = TRUE,
 #   scrollY = 200,
 #   scroller = TRUE ))
```


## Plot
```{r}
g <- ggplot(playersdf, aes(x=playersdf$pre_rating, y=playersdf$total_pts)) + geom_point() + geom_smooth(method="lm")
```

```{r}
g + ggtitle("Total Points vs Players' Pre-Rating", subtitle="From Chess Tournament Dataset") + xlab("Pre-Rating") + ylab("Total")
```


## Export to .CSV
```{r}
write.table(playersdf, file = "chessdata.csv", sep = ",", row.names = FALSE)
```

